---
title: "Question 3"
author: "Karen Abraham"
date: "12-09-2021"
output:
  html_document:
    df_print: paged
---
# Question 3: Which actors have been involved in the most conflicts since 1989?

## Loading Data and Libraries

```{r}
#clear environment
rm(list = ls())
options(warn=-1)

#loading libraries (must be installed before where necessary)
library(readr)
library(dplyr)
library(ggplot2)
library(network)
library(igraph)
library(centiserve)
library(tidygraph)
library(networkD3)
library(visNetwork)
library(htmlwidgets)
library(ggraph)
```

```{r}
#Loading the dataset
ged211 <- read_csv("ged211.csv")
head(ged211)
```

# Network Analysis
## Creating the network

First we need a usable dataframe
```{r}
#Lets take each instance with a death toll over 0 and create a df with only side_a, side_b and total death toll
#remove type of violence =3 because civilians are not considered actors
over0 <- subset(ged211, best>0 & type_of_violence!=3)
over0 <- cbind.data.frame(over0$side_a,over0$side_b,over0$best)
colnames(over0) <- c("side_a","side_b","best")

#Grouping and totalling
gpd <- aggregate(over0$best, list(over0$side_a,over0$side_b), FUN=sum) 
colnames(gpd) <- c("side_a","side_b","total")
gpd <- gpd %>% arrange(desc(gpd$total))

#Look at the top values
head(gpd)
```

Then we need to get all the nodes, which are actors listed in either side_a or side_b

```{r}

#All actors listed in side_a
sides_a <- gpd %>%
  distinct(side_a)

#All actors listed in side_b
sides_b <- gpd %>%
  distinct(side_b)

colnames(sides_a) <- c("sides")
colnames(sides_b) <- c("sides")

#Defining the nodes
nodes <- full_join(sides_a, sides_b, by = "sides")

#Defining the edges
edges <- gpd %>% 
  left_join(nodes, by = c("side_a" = "sides"))

edges <- edges %>% 
  left_join(nodes, by = c("side_b" = "sides")) 

```

Now we can create the network
```{r}
#Bidirectional so multiple set to true
routes_network <- network(edges, vertex.attr = nodes, matrix.type = "edgelist",ignore.eval = FALSE, multiple=TRUE)

#Checking we have created a network
class(routes_network)
routes_network
```

Now the network can be plotted
```{r}
plot(routes_network, vertex.cex = 0.8, main="Network Structure")
```

Although this gives us a little bit of insight into the structure, it is not very meaningful by itself, so let's work with an igraph

```{r}
routes_igraph <- graph_from_data_frame(d = edges, vertices = nodes, directed = TRUE)
routes_igraph
```

Using this network, we can easily query information on whether two actors are in conflict
```{r}
are.connected(routes_igraph,"Government of India","Government of Pakistan") #TRUE
are.connected(routes_igraph,"Government of Syria","IS") #TRUE
are.connected(routes_igraph,"Government of Syria","Government of Pakistan") #FALSE
```
We can also display the network as an edge list for each actor
```{r}
(as_adj_edge_list(routes_igraph))$`Government of Syria`
```

```{r}
routes_network
```

```{r}
V(routes_igraph)$all <- degree(routes_igraph, mode = "all")

#Which actors have more than 10 times the median number of edges (most interconnected vertices)?
V(routes_igraph)[V(routes_igraph)$all > 10*median(V(routes_igraph)$all)]
```


## Degree Distribution 

What does the distribution of edges look like
```{r}
#Minimum number of edges in network
min(degree(routes_igraph))

#Maximum number of edges in network
max(degree(routes_igraph))
```

```{r}
#Boxplot with distr of edges
boxplot(V(routes_igraph)$all, outline=FALSE, main="Boxplot of Edges")
```

```{r}
#Vertex degree distr
hist(V(routes_igraph)$all, breaks=100, main="Degree Distribution of Organized Violence",xlab="# of Edges")
```


```{r}
degree_distribution(routes_igraph)
#For degree_distribution a numeric vector of the same length as the maximum degree plus one. The first element is the relative frequency zero degree vertices, the second vertices with degree one, etc.

pie(degree_distribution(routes_igraph),main="Number of Edges")

```

```{r}
#Number of vertices
gorder(routes_igraph)

#Number of edges
gsize(routes_igraph)

#Reading degrees of actors
head(degree(routes_igraph))
```

Now lets look at the actors that are in conflict with the most other actors
```{r}
#Dataframe with number of degrees, ordered
degs <- as.data.frame(degree(routes_igraph))
colnames(degs) <- c("Degree")
head(degs[order(-degs$Degree),])

#Highest degrees
most_degs <- subset(degs,degs>12)
most_degs
```
# Density 

```{r}
#Lets take each instance with a death toll over 0 and create a df with only side_a, side_b and total death toll
#remove type of violence =3 because civilians are not considered actors

#Three empty dataframes to be filled later
densities <- data.frame("Year" = 1989:2020, "Density" = NA)
latoras <- data.frame("Year" = 1989:2020, "Closeness" = NA)
meandeg <- data.frame("Year" = 1989:2020, "Degrees" = NA)
vertices_cnt <- data.frame("Year" = 1989:2020, "Nodes" = NA)
edges_cnt <- data.frame("Year" = 1989:2020, "Edges" = NA)

#Loop that calculates scores for the above dataframes for each year
#https://www.rdocumentation.org/packages/centiserve/versions/1.0.0
for (i in densities$Year) {
  
over0 <- subset(ged211, best>0 & type_of_violence!=3 & year==i)
over0 <- cbind.data.frame(over0$side_a,over0$side_b,over0$best)
colnames(over0) <- c("side_a","side_b","best")

#Grouping and totalling
gpd <- aggregate(over0$best, list(over0$side_a,over0$side_b), FUN=sum) 
colnames(gpd) <- c("side_a","side_b","total")
gpd <- gpd %>% arrange(desc(gpd$total))

#Look at the top values
head(gpd)


#All actors listed in side_a
sides_a <- gpd %>%
  distinct(side_a)

#All actors listed in side_b
sides_b <- gpd %>%
  distinct(side_b)

colnames(sides_a) <- c("sides")
colnames(sides_b) <- c("sides")

#Defining the nodes
nodes <- full_join(sides_a, sides_b, by = "sides")

#Defining the edges
edges <- gpd %>% 
  left_join(nodes, by = c("side_a" = "sides"))

edges <- edges %>% 
  left_join(nodes, by = c("side_b" = "sides")) 

routes_igraph <- graph_from_data_frame(d = edges, vertices = nodes, directed = TRUE)

densities$Density[densities$Year==i] <- edge_density(routes_igraph, loops = FALSE)
latoras$Closeness[latoras$Year==i] <- closeness.latora(routes_igraph)
meandeg$Degrees[meandeg$Year==i] <-mean(degree(routes_igraph))
vertices_cnt$Nodes[meandeg$Year==i] <-length(V(routes_igraph))
edges_cnt$Edges[meandeg$Year==i] <-length(E(routes_igraph))
}
densities
latoras
meandeg
plot(densities, main="Network Density over Time")
plot(latoras, main="Network Closeness over Time") #sum of inversed distances to all other nodes instead of the inversed of the sum of distances to all other nodes
plot(meandeg, main="Mean Degrees over Time")
plot(vertices_cnt, main="Node Count over Time")
plot(edges_cnt, main="Edge Count over Time")

```


```{r}
#Tidy and arrange

routes_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)
routes_igraph_tidy <- as_tbl_graph(routes_igraph)

routes_tidy %>% 
  activate(edges) %>% 
  arrange(desc(total))

routes_tidy

routes_tidy %>% 
  activate(nodes) %>% 
  mutate(degree = centrality_degree(mode = 'total'))
```

```{r}
routes_tidy %>% 
  activate(nodes)%>% 
  mutate(degree = centrality_degree(mode = 'total'))  %>% 
  ggraph('nicely') + 
  geom_node_point(aes(size = degree)) +
  geom_edge_link()
```

```{r}
routes_tidy %>% 
    mutate(centrality = centrality_authority()) %>% 
    ggraph(layout = 'kk') + 
    geom_edge_link() + 
    geom_node_point(aes(colour = centrality)) + 
    scale_color_continuous(guide = 'legend') + 
    theme_graph()
```

As we can see, the edges are not weighted

```{r}
is.weighted(routes_igraph)
```

So let's weight by the total deaths in for conflict

```{r}
#Tidy and arrange
#tidygraph
routes_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)
routes_igraph_tidy <- as_tbl_graph(routes_igraph)

routes_tidy %>% 
  activate(edges) %>% 
  arrange(desc(total))

routes_tidy %>% 
  activate(nodes) %>% 
  mutate(degree = centrality_degree(mode = 'total'))
```


```{r}
head(E(routes_tidy)$total)

routes_tidy <- set_edge_attr(routes_tidy, "weight", value= E(routes_tidy)$total)
is_weighted(routes_tidy)

E(routes_tidy)[E(routes_tidy)$weight>0]

nodes[c(652,130,6,14,240,14,653,654,37,285,655,656,657),]

nodes[1:10,]
```

```{r}
#Fraction of deaths due to actors in most conflicts and deadliest conflicts

top10actors <- subset(E(routes_tidy), V(routes_tidy)$sides %in% c("IS","Fulani","Government of India","Government of Chad","Government of Myanmar (Burma)","Government of DR Congo (Zaire)", "Government of Sudan","Government of Pakistan","Government of Mali","Government of Ethiopia"))

sum(top10actors$total)/(sum(E(routes_tidy)$total))

#10 deadliest dyads
sum(E(routes_tidy)[1:10]$total)/(sum(E(routes_tidy)$total))

pie(top10actors$total)
pie(E(routes_tidy)[1:10]$total)
```

```{r}
head(closeness.latora(routes_tidy, vids = V(routes_tidy))) #Latora centrality for disconnected graphs

```

Using this we can see what the distribution of closeness looks like
```{r}
boxplot(closeness.latora(routes_tidy, vids = V(routes_tidy)), main="Closeness Distribution")
```
```{r}
#library(ggraph)
#ggraph(routes_tidy, layout = "graphopt") +   geom_node_point() +  geom_edge_link(aes(width = "weight"), alpha = 0.8) +   scale_edge_width(range = c(0.1, 3)) +  geom_node_text(aes(label = sides), repel = TRUE) +  labs(edge_width = "weight") +  theme_graph()
```


Interactive Plot with visNetwork and D3forcenetworklibrary

Jesse Sadler. (2017, October 25). Introduction to Network Analysis with R. Retrieved from https://www.jessesadler.com/post/network-analysis-with-r/

Bradley, B. (2021, May 25). From igraph to visNetwork. Retrieved from https://towardsdatascience.com/from-igraph-to-visnetwork-7bc5a76fdeec

```{r}
visIgraph(routes_igraph)
```

```{r}
visIgraph(routes_igraph) %>% visOptions(nodesIdSelection = TRUE,  highlightNearest = TRUE)
```


```{r}
library(networkD3)
library(visNetwork)

#This finds the subclusters
sc <- cluster_walktrap(routes_igraph)
members <- membership(sc)

#Converting from igraph to a d3 network type
plt_d3 <- igraph_to_networkD3(routes_igraph, group = members)

#This creates the plot
pltint <- forceNetwork(Links = plt_d3$links, Nodes =plt_d3$nodes,
             Source = 'source', Target = 'target', NodeID = 'name',Group = 'group', fontSize = 30, opacity = 1, zoom = TRUE)

pltint

#This saves the plot as an html file
saveWidget(pltint, "network.html")
```


## Libraries Used
### Documentation and functionality

readr: https://www.rdocumentation.org/packages/readr/versions/1.3.1
dplyr: https://www.rdocumentation.org/packages/dplyr/versions/0.7.8
ggplot2: https://ggplot2.tidyverse.org/

Inferring and Analysing the Network
network: https://cran.r-project.org/web/packages/network/index.html
igraph: https://igraph.org/r/
centiserve: https://www.centiserver.org/rpackage/
tidygraph: https://www.rdocumentation.org/packages/tidygraph/versions/1.2.0

Visualising the Network
networkD3: https://www.rdocumentation.org/packages/networkD3/versions/0.4
visNetwork: https://www.rdocumentation.org/packages/visNetwork/versions/2.1.0
htmlwidgets: https://www.htmlwidgets.org/
ggraph: https://cran.r-project.org/web/packages/ggraph/index.html

## Sources Used
### For the interactive plots:
Jesse Sadler. (2017, October 25). Introduction to Network Analysis with R. Retrieved from https://www.jessesadler.com/post/network-analysis-with-r/

Bradley, B. (2021, May 25). From igraph to visNetwork. Retrieved from https://towardsdatascience.com/from-igraph-to-visnetwork-7bc5a76fdeec
